Probe or Wait: Handling tail losses using Multipath TCP

### 2017-03-13 main.tex
Update

\begin{abstract}
Packet losses are known to affect the performance of latency sensitive applications in the Internet such as media streaming and gaming. Transport protocols recover from packet loss in order to provide reliable end to end communication and improving the quality of user experience. The efficiency of loss recovery greatly influences the completion time of flows. In this paper we focus on the state of the art loss recovery mechanisms for TCP and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms by calculating the burst completion time for each scenario. Based on the observations, we propose an enhancement to the tail loss recovery in Multipath TCP to improve the loss recovery time. Our experiment results show end to end latency performance improvement in controlled scenarios by reducing the burst completion time. 
\end{abstract}

@@

TCP has two mechanisms for detecting and recovering from packet losses: Fast Retransmit (FR) and Retransmission Timeout~(RTO)~\cite{Flach:2013}. Fast retransmit is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as a loss indication. In case of RTO, a sender waits for the packet loss to trigger a timeout before retransmitting that packet. During timeout, the sender cannot send any packets, but with FR the out-of-order packets trigger duplicate acknowledgements. The sender can start retransmission of missing packets without waiting for timeout. For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and recovering packet losses. Fast retransmit coupled with fast recovery can send data outstanding whene there is enough data to send in long flows. On the other hand, shorter flows more frequently depend on RTO for packet loss detection and recovery, thus incurring more latency. A single packet loss in a short flow may take many RTTs to detect and recover. This scenario is also applicable to the packets at the end of the flow (or tail), in a long flow. Although, RTO is reliable, even in tail loss~(losses at the end of transmission) recovery, it is not efficient in terms of recovery completion time.

@@

\section{Background and Related Work}\label{relwork}

Retransmission schemes are of significant importance in achieving low flow completion time. Several research studies has led to improved retransmission strategies for TCP. The current Linux TCP implementation provides room for improvement in several packet loss scenarios.

@@

The Linux implementation of MPTCP uses a set of retransmission heuristics to handle retransmissions. The data outstanding on a timed out subflow should be rescheduled for transmission on a different subflow using timeout as the indicator. Fast retransmit on a subflow will not trigger retransmission on another subflow. The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 4G vs WLAN, leading to flows with different delay characteristics. Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fastest paths. Such quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always to follow TCP semantics. MPTCP standard~\cite{rfc6824} suggests a more conservative approach. MPTCP is designed to exploit the multiple interfaces and often one or more of these could be wireless with completely different characteristics than that of wired interfaces. Authors in~\cite{Shin} argue that the calculation of RTO for MPTCP flows should include the interface characteristics to improve loss recovery time.

@@

\section{Experimental Setup}\label{exsetup}

Experiments use the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}
with the client node connected to two wireless interfaces 3G or 4G and WLAN and the server node connected to a wired router.
CORE uses the Linux networking stack in each node of the topology by using lightweight virtualization technique known as Linux network namespaces.
Characteristics of the connection setup and assumptions about the parameters are provided in Table~\ref{tab1}.
The parameters are chosen to be close to real values of bandwidth and delay of respective technologies.
There is no attempt in this study to focus on the effect of link detailed characteristics in retransmission performance.

@@

      \textbf{RTT} & \multicolumn{10}{c|}{20ms-120ms(3G/4G/WLAN)}  \\
      \hline 	
      \textbf{Bandwidth} & \multicolumn{10}{c|}{54Mbps(3G/4G/WLAN)}  \\
      \hline
      \textbf{Loss Model} & \multicolumn{10}{c|}{Deterministic}\\
      \hline

@@

\subsection{Testing retransmission with deterministic loss patterns}
In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
Losses are generated by associating netem with corresponding interfaces and dropping select packets. This process is simplified by using the KAUNetem tool~\cite{Garcia2016}, which allows experimenter to e.g., drop packets based on their positions in a flow. MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between the client and the server. So one has to wait both in time and packets for the second subflow establishment. We send two bursts of 80 packets each and drop tail packets in the second burst at one of the interfaces. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that subflow. We calculate the burst completion time for second burst that has specified tail losses. Three test cases are evaluated with one packet tail loss, two packet tail loss and one packet along with probe~(in TLP) or retransmission~(in RTO) loss.

@@

The setup has server and client with Linux supporting MPTCP running on them. Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms. The intention is to evaluate the effect of symmetry, mild asymmetry and higher aysmmetry in the link delays. We consider 5 scenarios with delay pairings 20ms-20ms, 20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance as shown in Table~\ref{tab1}.

\section{Observations and Discussion}\label{disc}
This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, retransmission or probe loss and two packet 
tail loss.

\subsection{TCP}
Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
starts with the results using TCP and compare it with that of MPTCP. Figures ~\ref{t1p},~\ref{t2p},~\ref{t1pp} represent the comparison of TCP 
burst completion times. Path asymmetry is irrelevant in this case as TCP uses a single path.

In the case of single packet tail loss, there is no difference in burst completion times of RTO and TLP settings. The retransmission
timeout and the probe timeout are same as there is only one outstanding packet. This is a special case in the TLP
specification that adds 200~ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts
for the delayed ACK and waits 200~ms before sending a probe. The minimum of PTO and RTO is considered for PTO.

@@

\begin{figure*}[!tbp]
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T1P.pdf}
	\caption{Single packet}\label{t1p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
	\caption{Two packet }\label{t2p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
	\caption{Single packet and retransmission }\label{t1pp}
 \end{subfigure}
 \caption{Tail loss scenarios using TCP}
\end{figure*}

\subsection{MPTCP}

The expected retransmission behavior of MPTCP for RTO and Linux default settings is shown 
in Figures.~\ref{timing1P},~\ref{timing1PP} and ~\ref{timing2P}. The actual pattern might differ in cases,
where the RTT is larger than 200~ms.

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing1P.pdf}
\end{center}
\caption{Timing diagram of MPTCP behavior with one packet tail loss a) RTO b) TLP}\label{timing1P}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing2P.pdf}
\end{center}
\caption{Timing diagram of MPTCP behavior with two packet tail loss a) RTO b) TLP}\label{timing2P}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing1PP.pdf}
\end{center}
\caption{Timing diagram of MPTCP behavior with one packet and next probe loss a) RTO b) TLP}\label{timing1PP}
\end{figure}

MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
the data until its congestion window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance
is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP for 
cases when the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
scenarios~(30ms-20ms,120ms-20ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
in Figure~\ref{timing1P}.

\begin{figure*}[!tbp]
 \begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/1P.pdf}
\caption{Single packet}\label{1p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/2P.pdf}
\caption{Two packet}\label{2p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/1PP.pdf}
\caption{Single packet and probe loss}\label{1pp}
 \end{subfigure}
 \caption{Tail loss scenarios using MPTCP}
\end{figure*}

@@

\section{Improvements to TLP for MPTCP}\label{impr}
The goal of Tail Loss Probe~(TLP) is to reduce tail latency of short flows. It achieves this by converting retransmission timeouts (RTOs) occurring due to tail losses into fast recovery. TLP transmits one packet in two round-trips when a connection is in Open state and isn't receiving any ACKs. The transmitted packet, aka loss probe, can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout. Our results show that the current TLP implementation for MPTCP helps improving the latency performance in cases when the paths are symmetric or the first path~(for the packet) has lowest RTT. However, it does not improve the performance in cases when the first path is not the lowest RTT path or when there is longer path interruption.

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.46\textwidth]{images/timingER3NewTLP.pdf}
\end{center}
\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) One packet b) Two packet c) One packet and Retransmission}\label{timingNew}
\end{figure}

\section{Evaluation of modified MPTCP TLP}\label{eval}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/1PNew.pdf}
\caption{Single packet tail loss comparison with MPTCP, TCP}\label{1pn}
\end{center}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/2PNew.pdf}
\caption{Two packet tail loss comparison with MPTCP, TCP}\label{2pn}
\end{center}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/1PPNew.pdf}
\caption{Single packet tail loss together with probe loss with MPTCP, TCP}\label{1ppn}
\end{center}
\end{figure}

\section{Conclusion}\label{conc}
This paper provides a short comparative analysis of retransmission behavior with TCP and MPTCP protocols using various loss patterns for tail loss.
The three cases of tail loss together with deterministic loss patterns allow us to dissect the retransmission behavior at packet level to understand possible improvements.
We propose a less conservative approach to trigger retransmission on an alternate path in the event of tail loss probe timeout, which otherwise triggered with an RTO. 
Our emulation experiments  using a modified Linux implementation, show that the proposed approach in fact improves the burst completion time in most cases and equals the existing
implementation in other cases. The proposed approach improves the performance when the alternate path has lower delay than the original loss path and the advantage increases with degree of asymmetry between the paths. Temporary path failures can cause probe loss along with packet loss. The approach is very effective in case of probe loss which otherwise incur RTO timeout to trigger retransmission on either path. However, the latency improvement comes at a cost of additional network traffic from the retransmitted packets. Further, we plan to investigate less conservative approaches for other retransmission events.

\bibliographystyle{IEEEtran}
\bibliography{mptcp-retrans}
\end{document}

>>>>

\begin{abstract}
Packet losses are known to affect the performance of latency sensitive applications in the Internet such as media streaming and gaming. Transport protocols recover from packet loss in order to provide reliable end to end communication and improving the quality of user experience. The efficiency of loss recovery greatly influences the completion time of flows. In this paper we focus on the state of the art loss recovery mechanisms for TCP and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms. Based on the observations, we propose an enhancement to the tail loss recovery in Multipath TCP to improve the loss recovery time. Our experiment results show consistent end to end latency performance improvement in considered scenarios. 
\end{abstract}

@@

TCP has two mechanisms for detecting and recovering from packet losses: Fast Retransmit (FR) and Retransmission Timeout~(RTO)~\cite{Flach:2013}. Fast retransmit is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as a loss indication. In case of RTO, a sender waits for the packet loss to trigger a timeout before retransmitting that packet. During timeout, the sender cannot send any packets, but with FR the out-of-order packets trigger duplicate acknowledgements. The sender can start retransmission of missing packets without waiting for timeout. For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and recovering packet losses. Fast retransmit coupled with fast recovery can send data outstanding when there is enough data to send in long flows. On the other hand, shorter flows more frequently depend on RTO for packet loss detection and recovery, thus incurring more latency. A single packet loss in a short flow may take many RTTs to detect and recover. This scenario is also applicable to the packets at the end of the flow (or tail), in a long flow. Although, RTO is reliable, even in tail loss~(losses at the end of transmission) recovery, it is not efficient in terms of recovery completion time.

@@

\kern-1em
\section{Background and Related Work}\label{relwork}

Retransmission schemes are of significant importance in achieving low flow completion time. Several IETF RFCs such as SACK based recovery~\cite{rfc6675}, Limited Transmit~\cite{rfc3042} and Early Retransmission~\cite{rfc5827} have led to improved retransmission strategies for TCP. TLP is part of the various deployable loss recovery mechanisms proposed in~\cite{Flach:2013}, suggesting that current Linux TCP implementation provides room for improvement in several packet loss scenarios.

@@

The Linux implementation of MPTCP uses a set of retransmission heuristics to handle retransmissions. The data outstanding on a timed out subflow should be rescheduled for transmission on a different subflow using timeout as the indicator. Fast retransmit on a subflow will not trigger retransmission on another subflow. The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 4G vs WLAN, leading to flows with different delay characteristics. Significant delay differences in available paths can cause receive buffer limitations, as the receiver has to wait for the data sent on slow path, while buffering the data on the fast path. In such cases, Opportunistic retransmission~\cite{Costin} improves the latency by using fast path to retransmit the data sent on an another path. Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fastest paths. Such quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always to follow TCP semantics. MPTCP standard~\cite{rfc6824} suggests a more conservative approach. MPTCP is designed to exploit the multiple interfaces and often one or more of these could be wireless with completely different characteristics than that of wired interfaces. Authors in~\cite{Shin} argue that the calculation of RTO for MPTCP flows should include the interface characteristics to improve loss recovery time.

@@

\kern-1em
\section{Experimental Setup}\label{exsetup}

Experiments use the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}
with the client node connected to two wireless interfaces 3G or 4G and WLAN and the server node connected to a wired router.
CORE uses the Linux networking stack in each node of the topology by using lightweight virtualization technique known as Linux network namespaces.
Characteristics of the connection setup and assumptions about the parameters are provided in Table~\ref{tab1}.
%The parameters are chosen to be close to real values of bandwidth and delay of respective technologies.
There is no attempt in this study to focus on the effect of link detailed characteristics in retransmission performance.

@@

      \textbf{RTT} & \multicolumn{10}{c|}{20\,ms-120\,ms(3G/4G/WLAN)}  \\
      \hline 	
      \textbf{Bandwidth} & \multicolumn{10}{c|}{54Mbps(3G/4G/WLAN)}  \\
      \hline
      \textbf{Loss Model} & \multicolumn{10}{c|}{Deterministic}\\
      \hline

@@

\kern-1em
\subsection{Testing retransmission with deterministic loss patterns}
In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
Losses are generated by associating netem with corresponding interfaces and dropping select packets. This process is simplified by using the KAUNetem tool~\cite{Garcia2016}, which allows experimenter to e.g., drop packets based on their positions in a flow. MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between the client and the server. So one has to wait both in time and packets for the second subflow establishment. We send two bursts of 80 packets each and drop tail packets in the second burst at one of the interfaces. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that subflow. We calculate the burst completion time for second burst that has specified tail losses. The objective of this experiment is to evaluate retransmission mechanisms with tail loss (single packet loss and multiple tail loss) and temporary path loss. Three test cases are considered for experiments: single packet tail loss and two packet tail loss for tail loss evaluation and single packet along with probe~(in TLP) or retransmission~(in RTO) loss for path loss evaluation.

@@

The setup has server and client with Linux supporting MPTCP running on them. Client has two wireless interfaces with one way delay on each interface ranging from 20\,ms to 120\,ms. The intention is to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in the link delays. We consider 5 scenarios with delay pairings 20\,ms-20\,ms, 20\,ms-30\,ms, 20\,ms-120\,ms, 30\,ms-20\,ms, 120\,ms-20\,ms to understand the effect of delay difference in the performance as shown in Table~\ref{tab1}.

\kern-1em
\section{Observations and Discussion}\label{disc}
This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, retransmission or probe loss and two packet 
tail loss.
\kern-5em
\subsection{TCP}
Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
starts with the results using TCP and compare it with that of MPTCP. Figures ~\ref{t1p},~\ref{t2p},~\ref{t1pp} represent the comparison of TCP 
burst completion times. Path asymmetry is irrelevant in this case as TCP uses a single path.

In the case of single packet tail loss, there is no difference in burst completion times of RTO and TLP settings. The retransmission
timeout and the probe timeout are same as there is only one outstanding packet. This is a special case in the TLP
specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts
for the delayed ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for PTO.

@@

\begin{figure*}[!tbp]
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T1P.pdf}
	\caption{Single packet}\label{t1p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
	\caption{Two packets }\label{t2p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
	\caption{Single packet and probe loss }\label{t1pp}
 \end{subfigure}
 \caption{Tail loss scenarios using TCP}
\end{figure*}
\kern-1em
\subsection{MPTCP}

The expected retransmission behavior of MPTCP for RTO and Linux default settings is shown 
in Figures.~\ref{timing1P},~\ref{timing1PP} and ~\ref{timing2P}. The actual pattern might differ in cases,
where the RTT is larger than 200\,ms.

\begin{figure*}[!tbp]
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing1P.pdf}
	\caption{Single packet}\label{timing1P}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth} 
	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing2P.pdf}
	\caption{Two packets }\label{timing2P}
 \end{subfigure} 
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
  	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing1PP.pdf}
	\caption{Single packet and probe loss}\label{timing1PP}
 \end{subfigure}
 \caption{Timing diagrams of MPTCP behavior on tail losses with a) RTO b) TLP }	
\end{figure*}

MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
the data until its congestion window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance
is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP for 
cases when both paths are symmetric or the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
in Figure~\ref{timing1P}.

\begin{figure*}[!tbp]
 \begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/1P.pdf}
\caption{Single packet}\label{1p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/2P.pdf}
\caption{Two packets}\label{2p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/1PP.pdf}
\caption{Single packet and probe loss}\label{1pp}
 \end{subfigure}
 \caption{Tail loss scenarios using MPTCP}
\end{figure*}

@@

\kern-1em
\section{Improvements to TLP for MPTCP}\label{impr}
The goal of Tail Loss Probe~(TLP) is to reduce tail latency of short flows. It achieves this by converting retransmission timeouts (RTOs) occurring due to tail losses into fast recovery. TLP transmits one packet in two round-trips when a connection is in Open state and isn't receiving any ACKs. The transmitted packet, aka loss probe, can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout. Our results show that the current TLP implementation for MPTCP helps improving the latency performance in cases when the paths are symmetric or the first path~(for the packet) has lowest RTT. However, it does not improve the performance in cases when the first path is not the lowest RTT path or when there is longer path interruption.

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.46\textwidth]{images/timingER3NewTLP1.pdf}
\end{center}
\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) Single packet b) Two packets c) Single packet and probe loss}\label{timingNew}
\end{figure}

\kern-1em
\section{Evaluation of modified MPTCP TLP}\label{eval}

@@

\begin{figure*}[!tbp]
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/1PNew.pdf}
\caption{Single packet}\label{1pn}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/2PNew.pdf}
\caption{Two packets}\label{2pn}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/1PPNew.pdf}
\caption{Single packet and probe loss}\label{1ppn}
\end{subfigure}
\caption{Performance comparison of MPTCP-New with MPTCP and TCP}
\end{figure*}

\kern-1em
\section{Conclusion}\label{conc}
This paper provides a short comparative analysis of retransmission behavior with TCP and MPTCP protocols using various loss patterns for tail loss.
The three cases of tail loss together with deterministic loss patterns allow us to dissect the retransmission behavior at packet level to understand possible improvements.
We propose a less conservative approach to trigger retransmission on an alternate path in the event of tail loss probe timeout, which otherwise triggered with an RTO. 
Our emulation experiments  using a modified Linux implementation, show that the proposed approach in fact improves the burst completion time in most cases and equals the existing
implementation in other cases. The proposed approach improves the performance when the alternate path has lower delay than the original loss path and the advantage increases with degree of asymmetry between the paths. Temporary path failures can cause probe loss along with packet loss. The approach is very effective in case of probe loss which otherwise incur RTO timeout to trigger retransmission on either path. However, the latency improvement comes at a cost of additional network traffic from the retransmitted packets. Further, we plan to investigate less conservative approaches for other retransmission events.

\kern-1em
\bibliographystyle{IEEEtran}
\bibliography{mptcp-retrans}
\end{document}
